Machine Learning Methods for Discourse Segmentation of Communications in E-Mail Based Behavioral Interventions

### 2018-07-21 paper-amia-summit-2018/manuscript.tex
update manuscript

\noindent{\bf Abstract}
\textit{Communication science approaches to developing effective behavior interventions, such as motivational interviewing (MI), are limited by traditionally manual qualitative coding of communication exchanges, which is a very resource-intensive and time-consuming process. This study focuses on the analysis of e-Coaching sessions, behavior interventions that are delivered via email and grounded in the principles of MI. A critical step towards automated annotation of e-Coaching communication exchanges is segmentation of emails into textual fragments that correspond to MI behaviors. In this work, we formulate this task as a classification problem and propose lexical, punctuation and topic features to address it. We experimented both with traditional machine learning methods, such as Support Vector Machine (SVM), Naive Bayes and K-Nearest Neighbor (KNN) classifiers and recurrent neural networks (RNNs). Results indicate that SVM outperformed KNN and RNNs achieving 0.990 macro F1-score overall and 0.845 and 0.995 macro F1-score for detecting ``new segment'' and ``same segment'' classes, respectively.}

@@

The goal of this research study is to assess the applicability of machine learning methods for automated segmentation of e-Coaching emails into textual fragments corresponding to individual behaviors, which is the first step of the coding process of e-Coaching communications. In particular, we introduced lexical, topic and punctuation features and experimented with both traditional supervised machine learning methods, such as Support Vector Machine (SVM), Naive Bayes (NB) and K-Nearest Neighbor (KNN) classifiers and deep learning methods, such as Long Short Term Memory (LSTM) and Gated Recurrent Unit (GRU), to find the best performing method and feature combination.

Relevant previous work in the biomedical domain primarily focused on segmentation of text into sections and headers\cite{apostolova2009automatic,denny2009evaluation,tepper2012statistical,cho2002text} or sentence boundary detection\cite{griffis2016quantitative,kreuzthaler2015detection,treviso2016sentence}. Apostolova et al.\cite{apostolova2009automatic} applied SVM along with word-vector cosine similarity metric combined with several heuristics to segment clinical reports into sections, such as demographics, history, procedure, finding and impression. After identification of each line in the document, Tepper et al. \cite{tepper2012statistical} trained Maximum Entropy models for section classification. Denny et al.\cite{denny2009evaluation} proposed a SecTag algorithm, which combined natural language processing techniques, terminology-based rules and Naive Bayes classifier to identify the sections and headers that achieved 99\% recall with 95.6\% precision. On the other hand, SVM based on prosodic and part of speech features \cite{kreuzthaler2015detection} and recurrent convolutional neural networks using word embeddings \cite{griffis2016quantitative} were utilized for detecting sentence boundaries. Segmentation of e-Coaching emails is different from a traditional shallow discourse analysis of conversations \cite{galley2003discourse} in that the focus is on segmentation, rather than on determining the types of transitions between the utterances or assigning utterances to speakers.

@@

\section*{Methods}
\subsection*{\textit{Data collection}}
The experimental dataset for this work was constructed from 49 e-coaching sessions, which include a total of 3,138 segmented and annotated MI behaviors. Each session represents an MI intervention delivered via email. During pre-processing, we removed all non-ASCII characters and applied stemming to normalize morphological variants of related concepts, such as ``eating'', ``eats'' and ``eat''. We formulate the segmentation task as a binary classification problem, as illustrated in Figure~\ref{fig:classifier} and experimented both with traditional machine learning methods, such as Support Vector Machine (SVM), Naive Bayes (NB) and K-Nearest Neighbor (KNN) classifiers and recurrent neural networks (RNNs), such as Long Short Term Memory (LSTM) and Gated Recurrent Unit (GRU). For NB, SVM and KNN models, e-Coaching email exchanges are partitioned into adjacent word pairs. Each pair is classified into either ``new segment'' or ``same segment'' class. The gold standard is manually segmented where adjacent word pairs are classified into ``new segment'' class. If all adjacent word pairs of a block of text are classified into ``same segment'' class, the entire block is treated as one textual fragment corresponding to a single MI behavior. In total, we obtained 95,421 word pairs, which include 3,138 ``new segment'' and 92,283 ``same segment'' instances. For RNNs, a block of text was taken as input sequence, such that one-hot encodings of each word or punctuation marks in a block were used as input into an RNN and binary labels (1 or 0) corresponding to ``new segment'' and ``same segment'' classification decision were considered as the output of RNN at each step. In the gold standard, words within the same segment were assigned the label of 0 and the last word or punctuation mark of a segment were assigned the label of 1.

\begin{figure}[!htb]
    \centering
    \includegraphics[width=0.80\textwidth]{figures/classifier.pdf}
    \caption{\textbf{Transformation of text segmentation task into text classification task}.}
    \label{fig:classifier}
\end{figure}

@@

We utilized three types of features in conjunction with SVM, NB and KNN methods: lexical, punctuation and topic features. Lexical features represent each word in a pair as a binary vector, in which 1 corresponds to the word in question and 0 to all other words. Since topic models\cite{kotov2015interpretable,hashimoto2016topic,lu2016modeling} have been shown to be effective semantic abstractions of individual words, we derived topic features from Labeled LDA\cite{ramage2009labeled}, a topic model for annotated corpora. We considered each textual segment in the training set as a document and the MI code assigned to the segment as a label. First, we derived a label-specific multinomial $p(w|l)$ for each label $l$ from the word distributions $p(w|l,z)$ for topic $z$ specific to label $l$ by marginalizing over topics:

\begin{equation}
\label{eq:pwl}
p(w|l) = {\sum_{z=1}^{K}{p(w|z,l)}}
\end{equation}

where $K$ is the optimal number of topics experimentally determined to minimize perplexity. After that, distributions over labels, $p(l|w)$, for each word were obtained from label-specific multinomials $p(w|l)$ using Bayesian inversion:

\begin{equation}
\label{eq:plw}
p(l|w) = \frac{p(w|l)p(l)}{p(w)}
\end{equation}

where $p(l)$ is the prior probabilities of label $l$ in the training set and $p(w)$ is the prior probability of word $w$ in the language model estimated from the training set using maximum likelihood. We use the distribution $p(l|w)$ as the topic feature vector, which shows how indicative each word $w$ is for a particular behavior code $l$. Topic features allow to abstract away from individual words and detect segment boundaries by indirectly capturing transitions between behavior codes when only observing a pair of words that are specific to different behavior codes. To find the optimal topic feature, we also derived the same topic feature $p(l|w)$ by using two recently proposed probabilistic generative latent variable models: Discriminative Labeled Latent Dirichlet Allocation (DL-LDA) and Latent Class Allocation (LCA)\cite{kotov2015interpretable}. Similar as L-LDA, DL-LDA-based topic feature is extracted by first deriving a label-specific language model $p(w|l)$ for each label $l$ from the word distributions $p(w|l,z)$ for topic $z$ specific to label $l$ by using the Eq.~\ref{eq:pwl} and \ref{eq:plw}. On the other hand, LCA directly provides label-specific multinomials $p(w|l)$, which can be used to derive topic feature by using Eq.~\ref{eq:plw}. Punctuation marks, which correspond to one of the symbols \{`.', `,', `!', `?', `:', `;', `-'\} between a pair of words, are also employed as a feature since punctuation marks designate the boundary of a sentence, clause and phrase and often also correspond to segment boundary.

@@

We experimented both with traditional classifiers, including Naive Bayes (NB)\cite{mccallum1998comparison}, Support Vector Machine (SVM)\cite{cortes1995support}, K-Nearest Neighbor (KNN)\cite{keller1985fuzzy} and two variants of Recurrent Neural Networks (RNNs)\cite{lipton2015critical}: Long Short Term Memory (LSTM)\cite{hochreiter1997long} and Gated Recurrent Unit (GRU)\cite{chung2014empirical}.

\textbf{Naive Bayes}: this model estimates the prior probability of classes along with conditional probabilities of each feature given the class. Then, the posterior probability is computed to predict the class for each sample by applying the Bayes rule with the assumption that features are conditionally independent. We used Multinomial Naive Bayes in experiments.

\textbf{Support Vector Machine}: state-of-the-art supervised classification method proven to be effective for text categorization\cite{joachims1998text} for its ability to cope with high dimensional input feature space. SVM finds the hyperplane that maximizes the separation between the closest ``new segment'' and ``same segment'' training examples. We scaled our dataset to utilize linear SVM with optimal cost 512 in experiments.

\textbf{K-Nearest Neighbor}: this method considers each training sample as a point in the input feature space. For a new test sample, Euclidean distance is calculated to find the k-nearest classified neighbors from the training set. Finally, the test sample is classified into a majority class of its k-nearest neighbors. We experimentally determined that best performance was achieved when the number of neighbors is 3.

\textbf{Recurrent Neural Networks}: RNN is a neural network architecture designed to capture sequential patterns. When we predict the new segment point using RNNs, particular combinations of words in a sequence will indicate the change of MI behavior. Long Short Term Memory (LSTM) networks\cite{hochreiter1997long} are a special type of RNN that are capable of handling variable size input sequence and have an internal memory that can be reset. GRU\cite{chung2014empirical} is a variant of LSTM mathematically represented by the following formula:

\begin{equation}
z_t = \sigma(W_zx_t + U_zh_{t-1} + b_z)
\label{eq:firstgru}
\end{equation}
\begin{equation}
r_t = \sigma(W_rx_t + U_rh_{t-1} + b_r)
\label{eq:resetgru}
\end{equation}
\begin{equation}
\tilde h_t = tanh(W_hx_t + r_t \odot U_hh_{t-1} + b_h) 
\label{eq:candidategru}
\end{equation}
\begin{equation}
h_t = z_t \odot h_{t-1} + (1-z_t) \odot \tilde h_t
\label{eq:lastgru}
\end{equation}  
In Eq.~\ref{eq:firstgru}-\ref{eq:lastgru}, $\sigma$ corresponds to sigmoid function and $\odot$ designates an element-wise product. The update gate $z_t$ and reset gate $r_t$ at time step $t$ are computed by the Eq.~(\ref{eq:firstgru}) and~(\ref{eq:resetgru}), where $W_z$, $W_r$, $W_h$, $U_z$, $U_r$, $U_h$ are the weight matrices and $b_z$, $b_h$ and $b_r$ are bias vectors. The activation $h_t$ of the GRU at time $t$ is a linear combination of previous activation $h_{t-1}$ and the candidate activation $\tilde h_t$, which is represented by Eq.~(\ref{eq:lastgru}) and~(\ref{eq:candidategru}). Our GRU model includes one hidden layer, output layer and input layer. We reset our model state after feeding each input sequence where input was given as one-hot encoding of word vector. Since LSTM and GRU use one-hot vector as input, results for these models are reported only when lexical and punctuation features are used. We experimentally determined that the best performance is achieved when the number of hidden units is 25, batch size is 1 and Adam\cite{kingma2014adam} is used for optimization. Source codes of this work will be available at https://github.com/teanalab/xxxxx upon acceptance of this paper for publication.

@@

\begin{table}[ht]
\centering
\caption{Performance of NB, SVM and KNN methods for identification of ``new segment'' and ``same segment'' classes as well as their weighted average. The highest value for each performance metric is highlighted in bold.}
\label{tab:result_base}
  \begin{tabular}{|l|l|l|l|l|l|l|l|l|l|}
  \hline
   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{new segment}} & \multicolumn{3}{|c|}{\textbf{same segment}}  & \multicolumn{3}{|c|}{\textbf{overall}} \\\cline{2-10}
   & \textbf{Precision}  & \textbf{Recall} & \textbf{F1} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1}\\ \hline    
 \multicolumn{10}{|c|}{lexical features only} \\ \hline   
 NB & 0.594 & 0.662 & 0.626 & 0.988 & 0.985 & 0.987 & 0.975 & 0.974 & 0.975  \\ \hline
 SVM & 0.721 & \textbf{0.705} & 0.713 & \textbf{0.990} & 0.991 & 0.990  & 0.981 & 0.981 & 0.981 \\ \hline
 KNN & \textbf{0.808} & 0.663 & \textbf{0.728} & 0.989 & \textbf{0.995} & \textbf{0.992} & \textbf{0.983} & \textbf{0.984} & \textbf{0.983} \\ \hline
 \multicolumn{10}{|c|}{lexical and puctuation features} \\ \hline
 LSTM & \textbf{0.810} & 0.720 & \textbf{0.762}  & \textbf{0.994} & \textbf{0.994} & \textbf{0.994} & \textbf{0.988} & \textbf{0.985} & \textbf{0.987} \\ \hline
 GRU & 0.783 & \textbf{0.739} & 0.760 & \textbf{0.994} & \textbf{0.994} & \textbf{0.994} & \textbf{0.988} & \textbf{0.985} & \textbf{0.987} \\ \hline
  \end{tabular}
\end{table}

@@

\begin{table}[ht]
\centering
\caption{Performance of NB, SVM and KNN methods with different topic model-based features for identification of ``new segment'' class in e-Coaching emails, when all features are used together. The highest value for each performance metric is highlighted in bold.}
\label{tab:result_boundary}
  \begin{tabular}{|l|l|l|l|l|l|l|l|l|l|}
  \hline
   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{L-LDA-based topic features}} & \multicolumn{3}{|c|}{\textbf{DL-LDA-based topic features}}  & \multicolumn{3}{|c|}{\textbf{LCA-based topic features}} \\\cline{2-10} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1}\\ \hline

 NB & 0.573 & 0.680 & 0.622 & 0.573 & 0.677 & 0.621 & 0.593 & 0.662 & 0.626 \\ \hline
 SVM & 0.831 & 0.783 & 0.806 & \textbf{0.877} & \textbf{0.815} & \textbf{0.845} & 0.867 & 0.804 & 0.834  \\ \hline
 KNN & 0.808 & 0.741 & 0.773 & 0.814 & 0.732 & 0.771 & 0.814 & 0.740 & 0.775 \\ \hline
  \end{tabular}
\end{table}

Table~\ref{tab:result_boundary} summarizes the performance of NB, SVM and KNN models for detecting ``new segment'' class. We observed that classification performance significantly improved for ``new segment'' class when lexical features are used in combination with punctuation and model-based topic features. Precision increases by 15.26\% and 0\%, 21.64\% and 0.74\% and 20.25\% and 0.74\% for SVM and KNN; recall increases by 11.06\% and 17.06\%, 15.60\% and 15.64\% and 14.04\% and 16.90\% for SVM and KNN; and F1-measure increases by 13.04\% and 6.18\%, 18.51\% and 5.91\% and 16.97\% and 6.46\% for SVM and KNN when L-LDA, DL-LDA and LCA-based topic features are used in conjunction combination with lexical and punctuation features, respectively. On the other hand, NB has very negligible impact of punctuation and model-based topic features. In contrast to results in Table~\ref{tab:result_base}, SVM demonstrates the best performance among all models achieving 0.877, 0.815 and 0.845 for precision, recall and F1-measure, respectively, when DL-LDA-based topic features are used in conjunction with lexical and punctuation features. NB and KNN provide their best performance when LCA-based topic features are used, achieving 0.626 and 0.775 F1-measure, respectively.\\

\begin{table}[ht]
\centering
\caption{Performance of NB, SVM and KNN methods with different topic model-based features for identification of ``same segment'' class in e-Coaching emails, when all features are used together. The highest value for each performance metric is highlighted in bold.}
\label{tab:result_not_boundary}
  \begin{tabular}{|l|l|l|l|l|l|l|l|l|l|}
  \hline
   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{L-LDA-based topic features}} & \multicolumn{3}{|c|}{\textbf{DL-LDA-based topic features}}  & \multicolumn{3}{|c|}{\textbf{LCA-based topic features}} \\\cline{2-10} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1}\\ \hline

 NB & 0.989 & 0.983 & 0.986 & 0.989 & 0.983 & 0.986 & 0.988 & 0.985 & 0.987 \\ \hline
 SVM & 0.993 & 0.995 & 0.994 & \textbf{0.994} & \textbf{0.996} & \textbf{0.995} & 0.993 & \textbf{0.996} & \textbf{0.995} \\ \hline
 KNN & 0.991 & 0.994 & 0.993 & 0.991 & 0.994 & 0.993 & 0.991 & 0.994 & 0.993 \\ \hline
  \end{tabular}
\end{table}

Similar to baseline performance, classification performance is remarkably higher for ``same segment'' class compared to ``new segment'' class (Table~\ref{tab:result_not_boundary}). The impact of punctuation and topic model-based features is also consistent with ``new segment'' classification while improvements are not as significant as ``new segment'' class. We observed that precision increases by 0.3\% and 0.2\%, 0.4\% and 0.2\% and 0.3\% and 0.2\% for SVM and KNN; recall increases by 0.4\% and -0.1\%, 0.5\% and -0.1\% and 0.5\% and -0.1\% for SVM and KNN; and F1-measure increases by 0.4\% and 0.1\%, 0.5\% and 0.1\% and 0.5\% and 0.1\% for SVM and KNN when L-LDA, DL-LDA and LCA-based topic features are used in conjunction combination with lexical and punctuation features, respectively. Similar to results in Table \ref{tab:result_boundary}, NB and KNN demonstrate the lowest and moderate performance on the task of identifying ``same segment'', respectively. \\

@@

@@

\section*{Conclusion}
Segmentation of e-Coaching emails is an integral part of developing e-Coaching interventions. Although several studies have focused on clinical interventions, they are limited by the qualitative coding of clinical interactions. In addition, previous studies in the medical domain mainly segmented clinical text into sections and sentences, none of them considered segmentation of text into groups of MI behaviors in the setting of discourse analysis with emails. In this paper, we compared the performance of machine learning models for the task of segmentation of e-Coaching text. We found out that SVM provides the best performance for the segmentation of text in terms of all performance metrics. Manual segmentation of e-Coaching data is very resource-intensive and time-consuming task, which can significantly decrease the time and effort required to develop an effective behavioral intervention. Our proposed methods can help to identify individual text segments, which can be annotated directly with a classification model. This approach will also help for developing fully automated e-Coaching and accelerate the pace of identifying effective communication strategies.

>>>>

\noindent{\bf Abstract}
\textit{Communication science approaches to developing effective behavior interventions, such as motivational interviewing (MI), are limited by the traditionally manual qualitative coding of communication exchanges, which is a very resource-intensive and time-consuming process. This study focuses on the analysis of e-Coaching sessions, behavior interventions that are delivered via email and grounded in the principles of MI. A critical step towards automated annotation of e-Coaching communication exchanges is segmentation of emails into textual fragments that correspond to MI behaviors. In this work, we formulate this task as a classification problem and propose lexical, punctuation and semantic features to address it. We experimented both with traditional machine learning method, Conditional Random Field (CRF) and deep learning methods, such as Multilayer Perceptrons (MLPs), Bidirectional Recurrent Neural Networks (BRNNs) and Convolutional Recurrent Neural Networks (CRNNs). Results indicate that CRNN outperformed CRF, MLP and BRNN achieving 0.990 macro F1-score overall and 0.821 macro F1-score for detecting new segment.}

@@

The goal of this research study is to assess the applicability of machine learning methods for automated segmentation of e-Coaching emails into textual fragments corresponding to individual behaviors, which is the first step of the coding process of e-Coaching communications. In particular, we introduced lexical, semantic and punctuation features and experimented with both traditional supervised machine learning method, linear-chain Conditional Random Field (CRF) and deep learning methods, such as Multilayer Perceptrons (MLPs), Bidirectional Recurrent Neural Networks (BRNNs) and Convolutional Recurrent Neural Networks (CRNNs), to find the best performing method and feature combination.

Relevant previous work in the biomedical domain primarily focused on segmentation of text into sections and headers\cite{apostolova2009automatic,denny2009evaluation,tepper2012statistical,cho2002text} or sentence boundary detection\cite{griffis2016quantitative,kreuzthaler2015detection,treviso2016sentence,fraser2015sentence}. Apostolova et al.\cite{apostolova2009automatic} applied SVM along with word-vector cosine similarity metric combined with several heuristics to segment clinical reports into sections, such as demographics, history, procedure, finding and impression. After identification of each line in the document, Tepper et al. \cite{tepper2012statistical} trained Maximum Entropy models for section classification. Denny et al.\cite{denny2009evaluation} proposed a SecTag algorithm, which combined natural language processing techniques, terminology-based rules and Naive Bayes classifier to identify the sections and headers that achieved 99\% recall with 95.6\% precision. On the other hand, SVM based on prosodic and part of speech features \cite{kreuzthaler2015detection} and recurrent convolutional neural networks using word embeddings \cite{griffis2016quantitative} were utilized for detecting sentence boundaries. Segmentation of e-Coaching emails is different from a traditional shallow discourse analysis of conversations \cite{galley2003discourse} in that the focus is on segmentation, rather than on determining the types of transitions between the utterances or assigning utterances to speakers.

@@

\section*{Methods}
\subsection*{\textit{Data collection}}
The experimental dataset for this work was constructed from 49 e-coaching sessions, which include a total of 3,138 segmented and annotated MI behaviors. Each session represents an MI intervention delivered via email. We formulate the segmentation task as a binary classification problem where each word annotated with a label 1 or 0, to indicate whether it preceds a new segment or not. We experimented both with traditional machine learning method, linear-chain Conditional Random Field (CRF) and deep learning methods, such as Multilayer Perceptrons (MLPs), Bidirectional Recurrent Neural Networks (BRNNs) and Convolutional Recurrent Neural Networks (CRNNs). For MLP models, e-Coaching email exchanges are partitioned into adjacent word pairs. Each pair is classified into either ``new segment'' or ``same segment'' class. The gold standard is manually segmented where adjacent word pairs are classified into ``new segment'' class. If all adjacent word pairs of a block of text are classified into ``same segment'' class, the entire block is treated as one textual fragment corresponding to a single MI behavior. In total, we obtained 95,421 word pairs, which include 3,138 ``new segment'' and 92,283 ``same segment'' instances. For deep learning models, an email was taken as input sequence, such that PoS tags and word embedding of each word or punctuation marks were used as input and binary labels (1 or 0) corresponding to ``new segment'' and ``same segment'' classification decision were considered as the output of the model at each step. In the gold standard, words within the same segment were assigned the label of 0 and the last word or punctuation mark of a segment were assigned the label of 1.

@@

We utilized three types of features in conjunction with CRF, MLP, BRNN and CRNN methods: lexical, punctuation and topic features. Lexical features represent each word in a pair as a binary vector, in which 1 corresponds to the word in question and 0 to all other words. Since part-of-speech (PoS) tags \cite{kotov2015interpretable,hashimoto2016topic,lu2016modeling} have been shown to be effective semantic abstractions of individual words, we used topic features for our experiment. Punctuation marks, which correspond to one of the symbols \{`.', `,', `!', `?', `:', `;', `-'\} between a pair of words, are also employed as a feature since punctuation marks designate the boundary of a sentence, clause and phrase and often also correspond to segment boundary.

@@

We experimented both with traditional classifiers, including Conditional Random Field (CRF) and deep learning models, such as Multilayer Perceptrons (MLPs) Bidirectional Recurrent Neural Networks (BRNNs) and Convolutional Recurrent Neural Networks (CRNNs).

\textbf{Conditional Random Field}: state-of-the-art supervised classification method proven to be effective for text categorization\cite{joachims1998text} for its ability to cope with high dimensional input feature space. SVM finds the hyperplane that maximizes the separation between the closest ``new segment'' and ``same segment'' training examples. We scaled our dataset to utilize linear SVM with optimal cost 512 in experiments.

\textbf{Multilayer Perceptrons}: this method considers each training sample as a point in the input feature space. For a new test sample, Euclidean distance is calculated to find the k-nearest classified neighbors from the training set. Finally, the test sample is classified into a majority class of its k-nearest neighbors. We experimentally determined that best performance was achieved when the number of neighbors is 3.

\textbf{Bidirectional Recurrent Neural Networks}: RNN is a neural network architecture designed to capture sequential patterns. When we predict the new segment point using RNNs, particular combinations of words in a sequence will indicate the change of MI behavior. Long Short Term Memory (LSTM) networks\cite{hochreiter1997long} are a special type of RNN that are capable of handling variable size input sequence and have an internal memory that can be reset. GRU\cite{chung2014empirical} is a variant of LSTM mathematically represented by the following formula:

\textbf{Convolutional Recurrent Neural Networks}: RNN is a neural network architecture designed to capture sequential patterns. When we predict the new segment point using RNNs, particular combinations of words in a sequence will indicate the change of MI behavior. Long Short Term Memory (LSTM) networks\cite{hochreiter1997long} are a special type of RNN that are capable of handling variable size input sequence and have an internal memory that can be reset. GRU\cite{chung2014empirical} is a variant of LSTM mathematically represented by the following formula:

Our GRU model includes one hidden layer, output layer and input layer. We reset our model state after feeding each input sequence where input was given as one-hot encoding of word vector. Since LSTM and GRU use one-hot vector as input, results for these models are reported only when lexical and punctuation features are used. We experimentally determined that the best performance is achieved when the number of hidden units is 25, batch size is 1 and Adam\cite{kingma2014adam} is used for optimization. Source codes of this work will be available at https://github.com/teanalab/xxxxx upon acceptance of this paper for publication.

@@

\begin{table}[ht]
\centering
\caption{Performance of NB, SVM and KNN methods for identification of ``new segment'' and ``same segment'' classes as well as their weighted average. The highest value for each performance metric is highlighted in bold.}
\label{tab:result_base}
  \begin{tabular}{|l|l|l|l|l|l|l|l|l|l|}
  \hline
   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{new segment}} & \multicolumn{3}{|c|}{\textbf{same segment}}  & \multicolumn{3}{|c|}{\textbf{overall}} \\\cline{2-10}
   & \textbf{Precision}  & \textbf{Recall} & \textbf{F1} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1}\\ \hline    
 \multicolumn{10}{|c|}{lexical features only} \\ \hline   
 NB & 0.594 & 0.662 & 0.626 & 0.988 & 0.985 & 0.987 & 0.975 & 0.974 & 0.975  \\ \hline
 SVM & 0.721 & \textbf{0.705} & 0.713 & \textbf{0.990} & 0.991 & 0.990  & 0.981 & 0.981 & 0.981 \\ \hline
 KNN & \textbf{0.808} & 0.663 & \textbf{0.728} & 0.989 & \textbf{0.995} & \textbf{0.992} & \textbf{0.983} & \textbf{0.984} & \textbf{0.983} \\ \hline
  \end{tabular}
\end{table}

@@

@@

\begin{figure}[!htb]
    \centering
    \includegraphics[width=1.0\textwidth]{figures/roc-curves.eps}
    \caption{\textbf{Receiver operating characteristic curves showing the performance of binary classifiers for the segmentation of e-Coaching text when lexical features (left) and combination of lexical, DL-LDA-based topic (except LSTM and GRU) and punctuation features (right) are used}.}
    \label{fig:roc-curves22}
\end{figure}

@@

\section*{Conclusion}
Segmentation of e-Coaching emails is an integral part of developing e-Coaching interventions. Although several studies have focused on clinical interventions, they are limited by the qualitative coding of clinical interactions. In addition, previous studies in the medical domain mainly segmented clinical text into sections and sentences, none of them considered segmentation of text into groups of MI behaviors in the setting of discourse analysis with emails. In this paper, we compared the performance of machine learning models for the task of segmentation of e-Coaching text. We found out that CRNN provides the best performance for the segmentation of text in terms of all performance metrics. Manual segmentation of e-Coaching data is very resource-intensive and time-consuming task, which can significantly decrease the time and effort required to develop an effective behavioral intervention. Our proposed methods can help to identify individual text segments, which can be annotated directly with a classification model. This approach will also help for developing fully automated e-Coaching and accelerate the pace of identifying effective communication strategies.
